Unsupervised Detection and Tracking of Arbitrary Objects with Dependent Dirichlet Process Mixtures

### 2012-03-23 ltx/paper_short_2.tex
added newStuff_ltx folder to hold lit review for new things; made some final edits on paper

\begin{abstract}
This paper proposes a technique for the unsupervised detection and tracking of arbitrary objects in videos. It is intended to reduce the need for detection and localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and image qualities. The technique uses a dependent Dirichlet process mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be easily and efficiently extracted from the regions in a video that represent objects. This paper describes a specific implementation of the model using spatial and color pixel data extracted via frame differencing and gives two algorithms for performing inference on the model to accomplish detection and tracking. This technique is demonstrated on multiple synthetic and benchmark video datasets that illustrate its ability to, without modification, detect and track objects with diverse physical charactersitics moving over non-uniform backgrounds and through occlusion.
% \keywords{Object Detection \and Tracking \and Bayesian Nonparametrics \and Multiple Target Tracking}
\end{abstract}

@@

We break-down this task into three parts: data extraction, localization, and tracking. Data extraction is the act of extracting features from regions of video-frames that constitute objects, localization is the act of finding the positions and/or shapes of distinct objects, and tracking is the act of maintaining the identities of the detected objects over time. This paper introduces a new framework for carrying out these three actions based on a type of dependent Dirichlet process mixture model. This framework provides a foundation for a class of unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.

Research related to general detection and tracking of objects tends to focus on one of either extraction, localization, or tracking. Integrating all three tasks in a system for multiple arbitrary objects and diverse video types is not often a primary focus. A few attempts at accomplishing the three tasks in a cohesive manner have been studied in recent years \cite{brostow2006unsupervised, brox2010object, fragkiadaki2011detection, pece_2002}.
%\willie{An example of general detection and tracking, called blob tracking, typically refers to methods that extract data and localize objects in each frame independently. Tracking involves finding blobs with similar locations and appearances over time. The main blob tracking challenges are the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from merging into a single object \cite{vermaak_2003}. These difficulties stem from performing detection via segmentation on individual frames.}
This paper furthers this line of work by providing a model that gives rise to a number of algorithms to detect arbitrary objects in videos---particularly in cases where frame-by-frame segmentation is difficult, video quality is low, and extraction is noisy---and maintain the isolation of distinct objects during tracking and through occlusion.

We begin by describing characteristics of the extracted data (Section~\ref{sec:dataextraction}), and giving the generalized form of the model (Section~\ref{sec:modeldefinition}). To implement this model, one must specify a data extraction procedure and distributions for representing objects, which may be chosen to allow for arbitrary object tracking or tailored to a specific object type for a given application. In our implementation, we extract data via a basic frame differencing procedure and specify distributions useful for representing arbitrary objects (Section~\ref{sec:modelspecification}). We describe inference algorithms for our model, and show how output of these algorithms can be interpreted as object localization and tracking results (Section~\ref{sec:inference}). Our implementation is demonstrated on multiple synthetic and benchmark datasets, and standard performance metric values are computed to quantify results on the benchmark datasets (Section~\ref{sec:experiments}). We compare our performance metrics with those yielded by specialized detection and tracking algorithms tailored to specific objects. Our results support our hypothesis that simple data extraction combined with inference on our model can perform detection and tracking of arbitrary objects at a level comparable to state-of-the-art, object specific algorithms.

% Prior Work
% ----------

@@

Methods that discern between the foreground and background regions of a video allow for data to be extracted from the areas in each frame where objects exist. Frame differencing and background subtraction are two such methods. Both record locations that exhibit motion relative to the background. Often, background subtraction refers to methods that compare an image containing targets with an image of the background only or with some model of the background that is learned as the video progresses \cite{piccardi2004background}, while frame differencing refers to methods that compare pairs of consecutive images in a video \cite{zhang2001segmentation}. Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking schemes \citep{perez_2002}.

A great deal of research has focused on developing algorithms to track multiple objects simultaneously. There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera) \cite{senior2006appearance, cucchiara2004probabilistic, zhou2003background}, complex object interactions \cite{khan_2004, mckenna2000tracking, dockstader2001multiple}, objects with similar appearances \cite{maccormick1999probabilistic, jepson_2003}, variable (and potentially high) numbers of objects \cite{reilly2010detection}, and objects that enter and exit a field of view at different times \cite{stauffer2003estimating, nedrich2010learning}. Multiple independent single-object trackers running simultaneously have been shown to be ineffective, as they will tend to coalesce and track the same object \cite{khan_2004}. To remedy this problem, methods such as \cite{maccormick1999probabilistic} have developed probabilistic principles for maintaining isolation of object trackers. An approach to this problem involving the use of a nonparametric mixture model has also found success in maintaining isolation of distinct objects \cite{vermaak_2003}.

Over the past decade, there have been attempts to provide general algorithms for the fully unsupervised detection and tracking of arbitrary objects in videos. Blob tracking, a basic method to carry out this goal, has found success in videos where objects are easily isolated from the background and where localization and segmentation of distinct objects is possible \cite{francois2004real, isard_2001}. Methods involving blob tracking, however, run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, and there exists object occlusion \cite{song2005model}. To aid the accuracy of these methods, techniques have been developed for performing extraction and segmentation in a joint manner, incorporating statistical methods for maintaining hypotheses of different numbers of detected objects, and introducing some of the multi-object tracking methods described previously to track distinct blobs after they have been segmented \cite{collins2003mean, isard_2001}. Another family of methods related to the task of unsupervised detection and tracking of video objects goes under the heading of video segmentation algorithms---these methods extend single-frame image segmentation to maintain coherence of image segments over time, and have had some success when used for the explicit purpose of detecting and tracking foreground objects in videos \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}. Other attempts to perform unsupervised detection and tracking include methods for clustering short sequences of positions extracted by detecting the motion of objects \cite{brostow2006unsupervised, brox2010object}, which aim to return full-length distinct object tracks, and graph based method that carry out a similar task using spectral clustering \cite{fragkiadaki2011detection}. Another approach uses a Gaussian mixture model to cluster data extracted from moving objects \cite{pece_2002}; this method also develops heuristics for the initialization and elimination of new object tracks.

Nearly all high accuracy object detection and tracking methods are those tailored for specific object types. These methods rely on ad-hoc detection criteria that exploit knowledge about the appearance or behavior of the given objects in a video. They often make use of state-of-the-art detectors designed to locate the specified objects of interest. In contrast, the method described in this paper is designed to track arbitrary objects without using any explicit detection criteria, and serve as a general strategy that can be used, without modification, to perform accurate detection and tracking of diverse objects in a wide range of videos. The method we introduce falls into the category of clustering-based arbitrary object detection and tracking methods. Differing from previous work, we use a type of time dependent Bayesian nonparametric mixture model, and show how it can be applied to a variety of easily extracted data to perform detection and tracking. This method begins by performing a simple data extraction procedure, which can be carried out broadly but yields noisy data. Our model of this data serves as a general framework for which we can choose a variety of object appearance distributions and inference algorithms; each choice provides a new method for unsupervised detection and tracking of arbitrary objects in videos.

@@

We desire a data extraction procedure that yields observations of the form
\begin{equation}
\centering
\bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
\end{equation}
where each $\bold{x}$ corresponds to a point within an image region where an object (or foreground element)  is believed to exist, $\bold{x}^{s} \in \mathbb{R}^{2}$ denotes the spatial location of this point, $\bold{x}^{c} \in S_{1} \times \ldots \times S_{V}$ denotes some collection of local image characteristics in the vicinity of this point, and $t \in \{1, \ldots, T \}$ denotes the time index.

We'd like to use an extraction procedure that is as unsophisticated as possible. Consequently, we use frame differencing. This procedure locates the pixels in image regions where there is motion relative to a previous frame. Here, each pixel corresponds to an observation $\bold{x}$. Frame differencing is simple, computationally inexpensive, and able to be applied to a wide range of static, single-camera videos (the videos used in experiments are stationary; moving-camera videos require data extraction methods useful for non-stationary video \cite{chien2002efficient, zhang2007moving}). Examples of pixel location data extracted via frame differencing are shown in Figure~\ref{fig:img_and_framediff}(a-j).

@@

% Model Definition
% --------------

@@

Dirichlet process mixture (DPM) models (Section \ref{sec:dpmixture}) fall under the heading of Bayesian nonparametric models. These models have been widely used in the past decade to perform nonparametric density estimation and cluster analysis. Data extracted from videos comprises spatiotemporal clusters, each corresponding with a distinct object. Consequently, we are interested in using a class of models known as dependent Dirichlet process mixture (DDPM) models (Section~\ref{sec:ddpmixture}), which are particularly useful for estimating the number of latent classes (clusters) in time dependent data. 
%In this work, estimating the number of clusters in video extraction data is equivalent to estimating the number of distinct objects in a video.

Since objects can enter and exit a scene, the number of clusters present throughout the video may not be constant (i.e., clusters may be created, or be ``born'', and may disappear, or ``die'', at intermediate time steps). To cluster data with these properties, we choose to use a DDPM known as the Generalized Polya Urn dependent Dirichlet process mixture, introduced by \cite{caron_2007}. This model may be intuitively viewed as a sequence of DPMs, where there exist dependencies between the parameters and number of clusters in adjacent time steps. Like the DPM, the GPUDDPM allows for the number of clusters within a dataset---which, in this work, corresponds to the number of objects in a video---to be found in an automated manner.

@@

In this model, each observation $\bold{x}_{i,t}$ is associated with an assignment variable $c_{i,t}$ that represents its assignment to a cluster $\theta_{k,t}$. 
%In a DPM, the probability of assignment to a cluster $k$ is a function of the cluster size $m_{k,t}$, which is equal to the number of observations that have been assigned to it. In contrast, 
The size (i.e. number of assigned observations) of each cluster in the GPUDDPM may decrease over time as observations become ``unassigned''. We define a distribution over the size of cluster $k$ at time $t$ ($m_{k,t}$), conditioned on the cluster's previous size ($m_{k,t-1}$), the assignments at time $t$ ($c_{1:N_{t}, t}$), and a deletion parameter ($\rho$),
\begin{equation}
\begin{split}
\label{D_distro}
\text{D} &(m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) := \hspace{2mm} m_{k,t-1} \\
&- \text{Binomial}( m_{k,t} | {m_{k,t-1}, \hspace{1mm} \rho}) + \sum_{i=1}^{N_{t}} \mathbb{I}(c_{i,t} = k)
\end{split}
\end{equation}
$\forall k \in \{1, \ldots, K_{t} \}$, where $K_{t}$ is the number of clusters at time $t$ and $\mathbb{I}(c_{i,t} = k)$ is an indicator function whose value is 1 if $c_{i,t} = k$ and 0 otherwise.

@@

Distributions \eqref{D_distro} and \eqref{C_distro} together comprise what \cite{caron_2007} refers to as the ``Generalized Polya Urn''. We can now define the GPUDDPM generatively as
\begin{align}
\label{gpuddpm_def}
\begin{split}
m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho  &\sim \text{D}(m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) \\
\theta_{k, t} | \theta_{k, t-1}   & \sim
\begin{cases}
  P(\theta_{k, t} | \theta_{k, t-1}) \hspace{1mm} \text{if} \hspace{1mm} k \leq K_{t} \\ % \in \{ 1, \ldots, K_{t} \} \\
  \mathbb{G}_{0} \hspace{1mm} \text{if} \hspace{1mm} k = K_{t+1}
\end{cases} \\
c_{i,t} | m_{1:K_{t},t}, \hspace{1mm} \alpha  &\sim  \text{C} (m_{1:K_{t},t}, \hspace{1mm} \alpha) \\
\bold{x}_{i,t} | c_{i,t}, \theta_{1:K_{t}, t} &\sim \text{F}(\theta_{c_{i,t}, t})
\end{split}
\end{align}
$\forall$ times $t \in \{1, \ldots, T\}$ and each cluster $k \in \{ 1, \ldots, K_{t} \}$ at time $t$, where we specify distributions for F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ in Section~\ref{sec:modelspecification}. The graphical model associated with this formulation is shown in Figure~\ref{fig:gpuddpm_gm_1}.
\begin{figure}[h]
        \center{\includegraphics[width=80mm]{../img/gpuddp_gm_1.pdf}}
        \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Polya Urn dependent Dirichlet process mixture. The observations at time $t$, $\bold{x}_{1:N_{t},t}$, and their associated assignments $c_{1:N_{t},t}$ are denoted respectively as $\bold{x}_{t}$ and $c_{t}$ (likewise for those at time $t-1$).}
\end{figure}

@@

\subsection{Object Appearance and Mixture Component F}
\label{sec:objectappearance}

At a given time $t$, we model each observation $\bold{x} \in \bold{X}$ as a draw from the product of a multivariate normal and multinomial distribution
\begin{equation}
\label{likelihood}
\text{F}(\bold{x}|\theta) = \mathcal{N}(\bold{x}^{s} | \boldsymbol{\mu}, \Sigma)  \mathcal{M}n(\bold{x}^{c} | \bold{p})
\end{equation}
where $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ denotes the parameters of a cluster at time $t$, with mean $\boldsymbol{\mu} \in \mathbb{R}^{2}$, covariance matrix $\Sigma \in \mathbb{R}^{2} \times \mathbb{R}^{2}$, and discrete probability vector $\bold{p} = (p_{1}, \ldots, p_{V})$ such that $\sum_{i=1}^{V}p_{i} = 1$. Additionally, $\mathcal{N}$ denotes the multivariate normal distribution and $\mathcal{M}n$ denotes the multinomial distribution. The mixture component F is sometimes referred to as the likelihood.

The multivariate normal distribution over the spatial features $\bold{x}^{s}$ models the position and spatial extent of an object. It can intuitively be thought to represent the shape of each object as an oval. We model the color features $\bold{x}^{c}$ as draws from a multinomial distribution. Incorporating this distribution into our cluster likelihood allows us to exploit our observation that the pixels associated with distinct objects tend to have similar color count vectors.

\subsection{Appearance Prior and Base Distribution $\mathbb{G}_{0}$}
\label{sec:appearanceprior}

$\mathbb{G}_{0}$ denotes the base distribution of the DDPM; it also serves as a prior distribution for the parameters $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ of the mixture components (i.e. of the object appearance distributions). We use conjugate priors in the base distribution to allow for more efficient computation. Specifically, in our implementation, a normal-inverse-Wishart prior is placed on the multivariate normal parameters $\{ \boldsymbol{\mu}, \Sigma \}$, and a Dirichlet prior is placed on the multinomial parameter $\bold{p}$ (where the normal-inverse-Wishart is a conjugate prior for the multivariate normal component of the likelihood and the Dirichlet is a conjugate prior for the multinomial component). The prior can therefore be written
\begin{equation}
\label{basedistro}
\mathbb{G}_{0}(\theta) = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}, \Sigma | \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0})  \mathcal{D}ir( \bold{p} | \bold{q}_{0})
\end{equation}
where $\mathcal{N}i\mathcal{W}$ denotes the normal-inverse-Wishart distribution, $\mathcal{D}ir$ denotes the Dirichlet distribution, and the prior has the hyperparameters $\boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0}$ and $\bold{q}_{0}$.

\subsection{Motion Model and Transition Kernel $P(\theta_{t} | \theta_{t-1})$}
\label{sec:motionmodel}

The transition kernel $P(\theta_{t} | \theta_{t-1})$ represents how we expect tracked objects to move over time. Since our implementation is intended for tracking arbitrary objects, we do not wish to make sophisticated assumptions about object motion. For example, we choose not to incorporate complex objects dynamics, though they are often used with success in certain object-specific tracking tasks, such as people tracking. We assume only that the position of an object at a given time is a function of its position at the previous time, and that the position varies in all directions equally between time steps.

>>>>

\begin{abstract}
This paper proposes a technique for the unsupervised detection and tracking of arbitrary objects in videos. It is intended to reduce the need for detection and localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and image qualities. The technique uses a dependent Dirichlet process mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be easily and efficiently extracted from the regions in a video that represent objects. This paper describes a specific implementation of the model using spatial and color pixel data extracted via frame differencing and gives two algorithms for performing inference in the model to accomplish detection and tracking. This technique is demonstrated on multiple synthetic and benchmark video datasets that illustrate its ability to, without modification, detect and track objects with diverse physical charactersitics moving over non-uniform backgrounds and through occlusion.
% \keywords{Object Detection \and Tracking \and Bayesian Nonparametrics \and Multiple Target Tracking}
\end{abstract}

@@

This task can be broken down into three parts: data extraction, localization, and tracking. Data extraction is the act of extracting features from regions of video-frames that constitute objects, localization is the act of finding the positions and/or shapes of distinct objects, and tracking is the act of maintaining the identities of the detected objects over time. This paper introduces a new framework for carrying out these three actions based on a type of dependent Dirichlet process mixture model. This framework provides a foundation for a class of unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.

Research related to general detection and tracking of objects tends to focus on one of either extraction, localization, or tracking. Integrating all three tasks in a system for multiple arbitrary objects and diverse video types is not often a primary focus. A few attempts at accomplishing the three tasks in a cohesive manner have been studied in recent years \cite{brostow2006unsupervised, brox2010object, fragkiadaki2011detection, pece_2002}.
%\willie{An example of general detection and tracking, called blob tracking, typically refers to methods that extract data and localize objects in each frame independently. Tracking involves finding blobs with similar locations and appearances over time. The main blob tracking challenges are the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from merging into a single object \cite{vermaak_2003}. These difficulties stem from performing detection via segmentation on individual frames.}
This paper furthers this line of work by providing a model that could give rise to a number of algorithms to detect arbitrary objects in videos---particularly in cases where frame-by-frame segmentation is difficult, video quality is low, and extraction is noisy---and maintain the isolation of distinct objects during tracking and through occlusion.

We begin by describing characteristics of the extracted data (Section~\ref{sec:dataextraction}), and giving the generalized form of the model (Section~\ref{sec:modeldefinition}). To implement this model, one must specify a data extraction procedure and distributions for representing objects, which may be chosen to allow for arbitrary object tracking or tailored to a specific object type for a given application. In our implementation, we extract data via a basic frame differencing procedure and specify distributions useful for representing arbitrary objects (Section~\ref{sec:modelspecification}). We describe inference algorithms for our model and show how the output of these algorithms can be interpreted as object localization and tracking results (Section~\ref{sec:inference}). Our implementation is demonstrated on multiple synthetic and benchmark datasets. Standard performance metric values are computed to quantify results on the benchmark datasets (Section~\ref{sec:experiments}). We compare our performance metrics with those yielded by specialized detection and tracking algorithms tailored to specific objects. Our results support our hypothesis that approaches combining simple data extraction and a powerful model can perform detection and tracking of arbitrary objects at a level comparable to state-of-the-art, object specific algorithms.

% Prior Work / Background
% -----------------------

@@

Methods that discern between the foreground and background regions of a video allow for data to be extracted from the areas in each frame where objects exist. Frame differencing and background subtraction are two such methods. Both record locations that exhibit motion relative to the background. Often, background subtraction refers to methods that compare an image containing targets with an image of the background only or with some model of the background that is learned as the video progresses \cite{piccardi2004background}, while frame differencing refers to methods that compare pairs of consecutive images in a video \cite{zhang2001segmentation}. Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking \citep{perez_2002}.

A great deal of research has focused on developing algorithms to track multiple objects simultaneously. There has been a particular emphasis on developing ways to deal with problems such as object occlusion (where one object blocks another from the view of a video camera) \cite{senior2006appearance, cucchiara2004probabilistic, zhou2003background}, complex object interactions \cite{khan_2004, mckenna2000tracking, dockstader2001multiple}, objects with similar appearances \cite{maccormick1999probabilistic, jepson_2003}, variable (and potentially high) numbers of objects \cite{reilly2010detection}, and objects that enter and exit a field of view at different times \cite{stauffer2003estimating, nedrich2010learning}. Multiple independent single-object trackers running simultaneously have been shown to be ineffective, as they will tend to coalesce and track the same object \cite{khan_2004}. To remedy this problem, methods have incorporated probabilistic principles for maintaining isolation of object trackers \cite{maccormick1999probabilistic}. An approach to this problem involving the use of a nonparametric mixture model has also found success in maintaining isolation of distinct objects \cite{vermaak_2003}.

Over the past decade, there have been attempts to provide general algorithms for the fully unsupervised detection and tracking of arbitrary objects in videos. Blob tracking, a basic method for carrying out this goal, has found success in videos where objects are easily isolated from the background and where localization and segmentation of distinct objects is possible \cite{francois2004real, isard_2001}. Blob tracking methods, however, run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, and there exists object occlusion \cite{song2005model}. To aid the accuracy of these methods, techniques have been developed for performing extraction and segmentation in a joint manner, incorporating statistical methods for maintaining hypotheses of different numbers of detected objects, and introducing some of the multi-object tracking methods described previously to track distinct blobs after they have been segmented \cite{collins2003mean, isard_2001}. Another family of methods related to the task of unsupervised detection and tracking of video objects goes under the heading of video segmentation algorithms---these methods extend single-frame image segmentation to maintain coherence of image segments over time, and have had some success when used for the explicit purpose of detecting and tracking foreground objects in videos \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}. Other attempts to perform unsupervised detection and tracking include methods for clustering short sequences of positions extracted by detecting the motion of objects \cite{brostow2006unsupervised, brox2010object}, which aim to return full-length distinct object tracks, and graph based methods that carry out a similar task using spectral clustering \cite{fragkiadaki2011detection}. Another approach uses a Gaussian mixture model to cluster data extracted from moving objects \cite{pece_2002}; this method also develops heuristics for the initialization and elimination of new object tracks.

Nearly all high accuracy object detection and tracking methods are those tailored for specific object types. These methods rely on detection criteria that exploit knowledge about the appearance or behavior of the objects in a video. Some of these methods make use of state-of-the-art detectors designed to locate the specified objects of interest. In contrast, the method described in this paper is designed to track arbitrary objects without using any explicit detection criteria, and serve as a general strategy that can be used, without modification, to perform accurate detection and tracking of diverse objects in a wide range of videos. The technique we introduce falls into the category of clustering-based arbitrary object detection and tracking methods. Differing from previous work, we use a type of time dependent Bayesian nonparametric mixture model, and show how it can be applied to a variety of easily extracted data to perform detection and tracking. This method begins by performing a simple data extraction procedure that yields noisy data. Our model of this data serves as a general framework for which we can choose a variety of object appearance distributions and inference algorithms; each choice provides a new method for unsupervised detection and tracking of arbitrary objects in videos.

@@

We desire a data extraction procedure that yields observations of the form
\begin{equation}
\centering
\bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
\end{equation}
where each $\bold{x}$ corresponds to a point within an image region where an object (or foreground element)  is believed to exist, $\bold{x}^{s} \in \mathbb{R}^{2}$ denotes the spatial location of this point, $\bold{x}^{c} \in S_{1} \times \ldots \times S_{V}$ denotes some collection of local image features in the vicinity of this point, and $t \in \{1, \ldots, T \}$ denotes the time index.

We'd like to use an extraction procedure that is as unsophisticated as possible. Consequently, we use frame differencing. This procedure locates the pixels in image regions that undergo change. Specifically, at each frame the pixels that differ from the previous frame beyond some threshold are recorded. Here, each pixel corresponds to an observation $\bold{x}$. Frame differencing is simple, computationally inexpensive, and able to be applied to a wide range of static, single-camera videos (the videos used in experiments are stationary; moving-camera videos require data extraction methods useful for non-stationary video \cite{chien2002efficient, zhang2007moving}). Examples of pixel location data extracted via frame differencing are shown in Figure~\ref{fig:img_and_framediff}(a-j).

@@

% Model Definition / Framework
% ----------------------------

@@

Dirichlet process mixture (DPM) models (Section \ref{sec:dpmixture}) fall under the heading of Bayesian nonparametric models. These models have been widely used in the past decade to perform nonparametric density estimation and cluster analysis. Data extracted from videos comprises spatiotemporal clusters, each corresponding to a distinct object. Consequently, we are interested in using a class of models known as dependent Dirichlet process mixture (DDPM) models (Section~\ref{sec:ddpmixture}), which are particularly useful for estimating the number of latent classes (clusters) in time dependent data. 
%In this work, estimating the number of clusters in video extraction data is equivalent to estimating the number of distinct objects in a video.

Since objects can enter and exit a scene, the number of clusters present throughout the video may not be constant (i.e., clusters may be created, or be ``born'', and may disappear, or ``die'', at intermediate time steps). To cluster data with these properties, we choose to use a DDPM known as the Generalized Polya Urn dependent Dirichlet process mixture \cite{caron_2007}. This model may be viewed intuitively as a sequence of DPMs, where there exist dependencies between the parameters and number of clusters in adjacent time steps. Like the DPM, the GPUDDPM allows for a distribution over the number of clusters within a dataset---which, in this work, corresponds to the number of objects in a video---to be inferred.

@@

In a GPUDDPM, each observation $\bold{x}_{i,t}$ is associated with an assignment variable $c_{i,t}$ that represents its assignment to a cluster $\theta_{k,t}$. 
%In a DPM, the probability of assignment to a cluster $k$ is a function of the cluster size $m_{k,t}$, which is equal to the number of observations that have been assigned to it. In contrast, 
The sizes of clusters in the GPUDDPM increase when observations are assigned to them, and decrease at later time points when these observations become ``unassigned''. We define a distribution over the size of cluster $k$ at time $t$ ($m_{k,t}$), conditioned on the cluster's previous size ($m_{k,t-1}$), the assignments at time $t$ ($c_{1:N_{t}, t}$), and a deletion parameter ($\rho$),
\begin{equation}
\begin{split}
\label{D_distro}
\text{D} &(m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) := \hspace{2mm} m_{k,t-1} \\
&- \text{Binomial}( m_{k,t} | {m_{k,t-1}, \hspace{1mm} \rho}) + \sum_{i=1}^{N_{t}} \mathbb{I}(c_{i,t} = k)
\end{split}
\end{equation}
$\forall k \in \{1, \ldots, K_{t} \}$, where $K_{t}$ is the number of clusters at time $t$ and $\mathbb{I}(c_{i,t} = k)$ is an indicator function whose value is 1 if $c_{i,t} = k$ and 0 otherwise.

@@

Distributions \eqref{D_distro} and \eqref{C_distro} together comprise what is referred to as the ``Generalized Polya Urn'' \cite{caron_2007}. We can now define the GPUDDPM generatively as
\begin{align}
\label{gpuddpm_def}
\begin{split}
m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho  &\sim \text{D}(m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) \\
\theta_{k, t} | \theta_{k, t-1}   & \sim
\begin{cases}
  P(\theta_{k, t} | \theta_{k, t-1}) \hspace{1mm} \text{if} \hspace{1mm} k \leq K_{t} \\ % \in \{ 1, \ldots, K_{t} \} \\
  \mathbb{G}_{0} \hspace{1mm} \text{if} \hspace{1mm} k = K_{t+1}
\end{cases} \\
c_{i,t} | m_{1:K_{t},t}, \hspace{1mm} \alpha  &\sim  \text{C} (m_{1:K_{t},t}, \hspace{1mm} \alpha) \\
\bold{x}_{i,t} | c_{i,t}, \theta_{1:K_{t}, t} &\sim \text{F}(\theta_{c_{i,t}, t})
\end{split}
\end{align}
$\forall$ times $t \in \{1, \ldots, T\}$ and each cluster $k \in \{ 1, \ldots, K_{t} \}$ at time $t$, where we choose application-specific distributions for F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ in Section~\ref{sec:modelspecification}. The graphical model associated with this formulation of the GPUDDPM is shown in Figure~\ref{fig:gpuddpm_gm_1}.
\begin{figure}[h]
        \center{\includegraphics[width=80mm]{../img/gpuddp_gm_1.pdf}}
        \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Polya Urn dependent Dirichlet process mixture. The observations at time $t$, $\bold{x}_{1:N_{t},t}$, and their associated assignments $c_{1:N_{t},t}$ are denoted respectively as $\bold{x}_{t}$ and $c_{t}$ (likewise for those at time $t-1$).}
\end{figure}

@@

\subsection{Object Appearance and Mixture Component, F}
\label{sec:objectappearance}

At a given time $t$, we model each observation $\bold{x} \in \bold{X}$ as a draw from the product of a multivariate normal and multinomial distribution
\begin{equation}
\label{likelihood}
\text{F}(\bold{x}|\theta) = \mathcal{N}(\bold{x}^{s} | \boldsymbol{\mu}, \Sigma)  \mathcal{M}n(\bold{x}^{c} | \bold{p})
\end{equation}
where $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ denotes the parameters of a cluster at time $t$, with mean $\boldsymbol{\mu} \in \mathbb{R}^{2}$, covariance matrix $\Sigma \in \mathbb{R}^{2\times2}$, and discrete probability vector $\bold{p} = (p_{1}, \ldots, p_{V})$ such that $\sum_{i=1}^{V}p_{i} = 1$. Additionally, $\mathcal{N}$ denotes the multivariate normal distribution and $\mathcal{M}n$ denotes the multinomial distribution. Note that we withhold writing the subscripts specifying the cluster number k and time t in this section when writing them is unnecessary.

The multivariate normal distribution over the spatial features $\bold{x}^{s}$ models the position and spatial extent of an object. It can be thought intuitively to represent the shape of each object as an oval. We model the color features $\bold{x}^{c}$ as draws from a multinomial distribution. Incorporating this distribution into our cluster likelihood allows us to exploit our observation that the pixels associated with distinct objects tend to have similar color count vectors.

\subsection{Appearance Prior and Base Distribution, $\mathbb{G}_{0}$}
\label{sec:appearanceprior}

$\mathbb{G}_{0}$ denotes the base distribution of the DDPM; it also serves as a prior distribution for the parameters $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ of the mixture components (i.e. of the object appearance distributions). We use conjugate priors in the base distribution to allow for more efficient computation. Specifically, in our implementation, a normal-inverse-Wishart prior is placed on the multivariate normal parameters $\{ \boldsymbol{\mu}, \Sigma \}$, and a Dirichlet prior is placed on the multinomial parameter $\bold{p}$. The prior can therefore be written
\begin{equation}
\label{basedistro}
\mathbb{G}_{0}(\theta) = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}, \Sigma | \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0})  \mathcal{D}ir( \bold{p} | \bold{q}_{0})
\end{equation}
where $\mathcal{N}i\mathcal{W}$ denotes the normal-inverse-Wishart distribution, $\mathcal{D}ir$ denotes the Dirichlet distribution, and the prior has the hyperparameters $\boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0}$ and $\bold{q}_{0}$.

\subsection{Motion Model and Transition Kernel, $P(\theta_{t} | \theta_{t-1})$}
\label{sec:motionmodel}

The transition kernel $P(\theta_{t} | \theta_{t-1})$ represents how we expect tracked objects to move over time. Since our implementation is intended for tracking arbitrary objects, we do not wish to make sophisticated assumptions about object motion. For example, we choose not to incorporate complex objects dynamics, though they are often used with success in certain object-specific tracking tasks, such as people tracking \cite{choo2001people, conte2010performance}. We assume only that the position of an object at a given time is close to its position at the previous time, and that the position varies in all directions equally between time steps.

The base distribution $\mathbb{G}_{0}$ must be the invariant distribution of the transition kernel $P(\theta_{t} | \theta_{t-1})$ in order for the the cluster parameters to remain marginally distributed according to the base distribution, and for the model to be a valid GPUDDPM. In other words, the transition kernel must satisfy
\begin{equation}
\int \mathbb{G}_{0}(\theta_{t-1})P(\theta_{t} | \theta_{t-1}) d\theta_{t-1} = \mathbb{G}_{0}(\theta_{t})
\end{equation}
for a given cluster with parameters $\theta$. One way to achieve this is through the use of auxiliary variables. These are a set of $M$ variables $\bold{z}_{t} = (z_{t,1}, \ldots, z_{t,M})$ associated with each cluster at each time $t$ that satisfy
\begin{eqnarray}
\label{auxiliaryvariablecriteria}
P(\theta_{t} | \theta_{t-1}) = \int P(\theta_{t} | \bold{z}_{t}) P(\bold{z}_{t} | \theta_{t-1}) d \bold{z}_{t}
\end{eqnarray}

With the addition of these variables, the parameters of a cluster at a given time do not depend directly on their value at the previous time; they are instead dependent through an intermediate sequence of variables. This allows the cluster parameters at each time step to be marginally distributed according to the base distribution $\mathbb{G}_{0}$ while maintaining simple time varying behavior.
